---
title: "IMDB Ratings: Does the director play a role?"
author: "Lasse Munk"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
library(kableExtra)
```
# About 
This project was done to research whether the IMDB rating for blockbuster movies were influenced by the director. Before looking at any kind of data, a null hypothesis and alternative hypothesis were set up as follows:

### Null Hypothesis
* H0 there is no difference in the mean IMDB-rating for movies produced by Steven Spielberg or Tim Burton. 

### Alternative Hypothesis
* Ha There is a significant difference between these mean IMDB-ratings.

## Research 
A big dataset was provided with movies from all genres by many different directors. To narrow down the research, just Steven Spielburg and Tim Burton will be compared. 

First I load the data into R: 

```{r load-movies-data}
movies <- read_csv(here::here("data", "movies.csv"))
glimpse(movies)
```

Then I manipulate the data set so that only Steven Spielburg and Tim Burton are evaluated. The standard error for their average IMDB movie rating is calculated to create confidence intervals. 

```{r}
movies1 <- movies %>% 
  filter(director %in% c("Steven Spielberg", "Tim Burton")) %>%
  select(director, rating) %>%
  group_by(director) %>%
  summarize(avg_rating= mean(rating), 
            count = n(),
            sderror = sd(rating)/sqrt(count-1)) %>%
  mutate(upper_limit = avg_rating + 1.96*sderror,
         lower_limit = avg_rating - 1.96*sderror) %>%
  arrange(avg_rating)
```

As we see from the data manipulation, the directors have the following statistics: 

```{r}
movies1 %>% 
  kbl() %>%
  kable_styling()
```
Steven Spielburg has an average IMDB rating of 7.57 while Tim Burton has an average rating of just 6.93. Furthermore, Steven has made more movies with a smaller spread across ratings, which is seen by his standard error of just 0.148. On the other hand, Tim Burton has made fewer movies with a larger discrepancy in ratings; causing him to have an standard error of 0.193. 

### Data Visualisation
Judging from these numbers alone, it does indeed seem that a movie's director has an impact on the IMDB Rating. But looking at numbers can be a bit daunting, so let's try to make this a bit more appealing to look at: 

```{r}
ggplot(movies1, mapping=aes(
  x = avg_rating,
  y= reorder(director, avg_rating), 
  colour = director)) + geom_errorbarh(
    movies1, 
    width=.03, 
  mapping=aes(
    xmin=upper_limit, 
    xmax=lower_limit)) + theme_bw() + 
  geom_point(movies1, mapping=aes(
    x=avg_rating, 
    size = 4)) + theme(legend.position = "none") + 
  geom_rect(
    movies1[1,], 
    mapping = aes(
      xmin = 7.28, 
      xmax = 7.31, 
      ymin = 0, 
      ymax = 3, 
      color = NA, 
      alpha = 0.2)) + 
  geom_text(aes(
    x = upper_limit, 
    y = director, 
    label=round(upper_limit, 2)), vjust = -1.5, color = "black") + 
  geom_text(aes(
    x = lower_limit, 
    y = director, 
    label=round(lower_limit, 2)), vjust = -1.5, color = "black") +
  geom_text(aes(
    x= avg_rating, 
    y = director, 
    label=round(avg_rating, 2)), vjust = -1.5, color = "black", size = 6) + 
  labs(title = "Do Spielberg and Burton have the same mean IMDB ratings?",            subtitle = "95% confidence intervals overlap") + 
  theme(axis.title.y = element_blank()) + 
  xlab("Mean IMDB Rating")
```
As the visualisation shows, the two directors' confidence intervals actually overlap by a very small margin. By convention, this usually means there is not a statistical significance, ie that the director does not play a determining factor in the IMDB rating. 

## Next steps - re-sampling the data
However, we have only sampled two directors out of a very large dataset - what if we choose two different ones? 

```{r}
movies2 <- movies %>% 
  filter(director %in% c("James Cameron", "Michael Bay")) %>%
  select(director, rating) %>%
  group_by(director) %>%
  summarize(avg_rating= mean(rating), 
            count = n(),
            sderror = sd(rating)/sqrt(count-1)) %>%
  mutate(upper_limit = avg_rating + 1.96*sderror,
         lower_limit = avg_rating - 1.96*sderror) %>%
  arrange(avg_rating)

movies2 %>%
  kbl() %>%
  kable_styling()
```
##  Conclusion
Using Michael Bay and James Cameron this becomes a lot more apparent - here the gap between the two confidence intervals is quite large and so **tells us that a movie's director most likely plays a role in its IMDB rating**. 
