---
title: "COVID Research"
author: "Lasse Munk"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, include=FALSE}
library(tidyverse)  
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(infer)
library(blogdown)
```

# About
The data in this research project is downloaded from the [CDC Covid-19 Case Surveillance Data](https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf).


```{r, cache=TRUE}
# URL link to CDC to download data
url <- "https://data.cdc.gov/api/views/vbim-akqf/rows.csv?accessType=DOWNLOAD"

covid_data <- vroom::vroom(url)%>%
  clean_names()
```

## Task 
Given the data we have, I set out to create two graphs showing the deaathrate in combination with

1.  age group, sex, and whether the patient had co-morbidities or not
2.  age group, sex, and whether the patient was admited to Intensive Care Unit (ICU) or not.

### 1. COVID and Co-morbidities 
First I create the graph for death rate and co-morbidities seperated by age group, sex and Y/N comorbidity. 

```{r}
graphdata_comorbidities <- covid_data %>%
  #Filter out patients with missing information regarding co-morbidity
  filter(medcond_yn %in% c("Yes", "No")) %>%
  #Rename Yes/No to With/Without comorbidities for graph representation
  mutate(medcond_yn = ifelse(medcond_yn == "Yes", "With comorbidities", "Without comorbidities")) %>%
  #Clean death rate column to only Yes/No  
  filter(death_yn %in% c("Yes", "No")) %>%
  #Filter out patients with N/A sex 
  filter(sex %in% c("Female", "Male")) %>%
  #Filter out unknown age groups
  filter(age_group != "Unknown") %>%
  #Narrow down the columns that are relevant
  select("sex", "age_group", "death_yn", "medcond_yn")

#Mutate the death rate into percentage for each age group 
graphdata_comorbidities <- graphdata_comorbidities %>% 
  #Group per age group
  group_by(age_group, sex, medcond_yn, death_yn) %>%
  #Aggregate numbers for selected categories
  count() %>%
  #Spread the values from column death_yn to enable us to calculate the          frequency
  spread(key = death_yn, value = n) %>%
  #Now we can calculate the frequency of cases per segment 
  summarize(deathrate_freq = Yes/(No+Yes))
  

#Plot creation, we use geom_cols because we want to represent values in the data 
comorbidity_plot <- ggplot(graphdata_comorbidities, aes(x = age_group, y = deathrate_freq)) + 
  geom_col(fill = "#3273a8", alpha = 0.7) + 
  #Round to 1 decimal point and turn into percentage, adjust towards right
  geom_text(aes(label = round(deathrate_freq*100, 1), hjust = -0.1)) +
  coord_flip() + 
  #Separate by sex and comorbidity
  facet_grid(vars(medcond_yn), vars(sex)) + 
  theme_bw() +
  labs(title = "Covid death % by age group, sex and presence of co-morbidities", caption = "Source: CDC") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())


comorbidity_plot
```
#### Conclusion
Here we see the following interesting conclusions: 

* Patients >=60 years old had a far higher mortality rate than other age groups 
* Patients with co-morbidities were far more likely to die as a result of COVID 
* Hardly anyone < 40 without comorbidities passed away from COVID 

### 2. COVID and admittance to ICU 
Let's now turn over attention to the death rate in combination with admittance to the ICU (Intensive Care Unit)
```{r}
graphdata_icu <- covid_data %>%
  #Filter out patients with missing information regarding icu admittance
  filter(icu_yn %in% c("Yes", "No")) %>%
  #Clean death rate column to only Yes/No  
  filter(death_yn %in% c("Yes", "No")) %>%
  mutate(icu_yn = ifelse(icu_yn == "Yes", "Admitted to ICU", "No ICU")) %>%
  #Filter out patients with N/A sex 
  filter(sex %in% c("Female", "Male")) %>%
  #Filter out unknown age groups
  filter(age_group != "Unknown") %>%
  #Narrow down the columns that are relevant
  select("sex", "age_group", "death_yn", "icu_yn")

#Mutate the death rate into percentage for each age group 
graphdata_icu <- graphdata_icu %>% 
  #Group per age group
  group_by(age_group, sex, icu_yn, death_yn) %>%
  #Aggregate numbers for selected categories
  count() %>%
  #Spread the values from column death_yn to enable us to calculate the          frequency
  spread(key = death_yn, value = n) %>%
  #Now we can calculate the frequency of cases per segment 
  summarize(deathrate_freq = Yes/(No+Yes))
  

#Plot creation, we use geom_cols because we want to represent values in the data 
icu_plot <- ggplot(graphdata_icu, aes(x = age_group, y = deathrate_freq)) + 
  geom_col(fill = "#f24f0f", alpha = 0.7) + 
  #Round to 1 decimal point and turn into percentage, adjust towards right
  geom_text(aes(label = round(deathrate_freq*100, 1), hjust = -0.1)) +
  coord_flip() + 
  #Separate by sex and comorbidity
  facet_grid(vars(icu_yn), vars(sex)) + 
  theme_bw() +
  labs(title = "Covid death % by age group, sex and whether patient was admitted to ICU", caption = "Source: CDC") + 
  scale_y_continuous(labels = scales::percent) + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

icu_plot
```
#### Conclusion 
Here we see the interesting conclusions:

* If a patient with COVID is above 59 years and is admitted to the ICU, there's a +50% chance he/she will die 
* Even for the most senior patients with COVID, non-admittance to the ICU greatly increases your survival odds 
* Admittance to the ICU is a far greater indicator of mortality than co-morbidity is 

A next iteration of this research would be to look at the relation between COVID-patients and whether co-morbidities were a contributing factor to being admitted to the ICU. 
